---
title: "COVID-19 in Poland"
description: |
  A short description of the post.
author:
  - name: Yalong Liu
    url: https://example.com/norajones
date: 09-01-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(DT)
library(sparkline)
```


```{r}
covid_raw <- read_csv(here::here("data/owid-covid-data.csv"))
```

# Data Description

```{r}
# Data wrangling
covid_tidy <- covid_raw %>% 
  filter(location == "Poland") %>% 
  select(c(date, total_cases, new_cases,
           total_deaths, new_deaths, total_cases_per_million, total_deaths_per_million, total_vaccinations,
           total_vaccinations_per_hundred, people_fully_vaccinated, 
           people_fully_vaccinated_per_hundred))
```



```{r}
desc <- data.frame(Variable = names(covid_tidy),
                   Type = c("date", "double", "double", "double", "double", "double", "double", "double", "double", "double", "double"),
                   Description = c("The observation date", "Total number of confirmed cases of COVID-19", 
                                   "The number of new cases per day", "Total number of deaths of COVID-19", 
                                   "Total number of confirmed cases per million", "Total number of deaths per million",
                                   "Total number of new deaths per day", "Cumulative total vaccine doses administered", 
                                   "Cumulative total vaccine doses administered per 100 population", "Cumulative number of persons fully vaccinated", 
                                   "Cumulative number of persons fully vaccinated per 100 population"
                                   ))
  knitr::kable(desc, bootabs = TRUE,
               table.attr="class='xtab'") %>% 
  kable_styling(bootstrap_options = "hover")
```


# Analysis and Exploration



## COVID-19 Cases in Poland

```{r}
cases_trend <- covid_tidy %>% 
  filter(date == max(date)) %>% 
  select(c(date, total_cases, total_cases_per_million, total_deaths,
           total_deaths_per_million)) %>% 
  mutate(total_cases = scales::comma(total_cases),
         total_deaths = scales::comma(total_deaths),
         total_deaths_per_million = round(total_deaths_per_million, 2))
  knitr::kable(cases_trend, bootabs = TRUE,
               caption = "Last situation of COVID-19 in Poland",
               align = "lrrrr") %>% 
  kable_styling() %>% 
  column_spec(2:5, width = "100px") %>% 
  add_header_above(c(" " = 1, "Cases" = 2, "Deaths" = 2)) %>% 
  kable_classic(full_width = FALSE)
```

```{r}
cases_month <- covid_tidy %>% 
mutate(month = format(as.Date(date), "%Y-%m")) %>% 
  select(month, new_cases, new_deaths) 

cases_mean <- cases_month %>% 
  group_by(month) %>% 
  summarise(new_cases_avg = round(mean(new_cases, na.rm = TRUE), 2),
            new_deaths_avg = round(mean(new_deaths, na.rm = TRUE), 2))

sparkline_data <- cases_month %>% 
  group_by(month) %>% 
  summarise(new_cases_line = spk_chr(new_cases, type = "line",
                                     chartRangeMin = 0, chartRangeMax = max(new_cases)),
            new_deaths_line = spk_chr(new_deaths, type = "line",
            chartRangeMin = 0, chartRangeMax = max(new_deaths)))

cases_trend <- left_join(cases_mean, sparkline_data, by = "month") %>% 
  separate(col = month,
  into = c("year", "month"), "-") %>% 
  mutate(year = as.double(year),
         month = as.double(month)) %>%
  select(c(year, month, new_cases_avg, new_cases_line, new_deaths_avg, new_deaths_line)) %>% 
  datatable(escape = FALSE, 
            filter = 'top',
            caption = htmltools::tags$caption("Monthly average statistics and monthly trends of new cases and new deaths in Poland"),
            colnames = c("Year", "Month", "New Cases (monthly_avg)", "New Cases Trend (monthly)",
                         "New Deaths (monthly_avg)", "New Deaths Trend (monthly)"),
            options = list(pageLength = 10,
                           columnDefs = list(list(className = 'dt-right', targets = 0:6)),
                           fnDrawCallback = htmlwidgets::JS(
  '
function(){
  HTMLWidgets.staticRender();
}
'
))) %>% 
  spk_add_deps()
cases_trend
```

### Numerical Summary Statistics

```{r}
num_stats <- covid_tidy %>% 
  select(new_cases, new_deaths) %>% 
  pivot_longer(cols = c(new_cases, new_deaths),
               names_to = "Variable",
               values_to = "Record") %>% 
  group_by(Variable) %>% 
  summarise(Mean = round(mean(Record, na.rm = TRUE),2),
            SD = round(sd(Record, na.rm = TRUE),2),
            Median = round(median(Record, na.rm = TRUE),2),
            Max = scales::comma(max(Record, na.rm = TRUE)))

  knitr::kable(num_stats, bootabs = TRUE,
               table.attr = "class='cor-table'",
               caption = "Numerical summary statistics of new coronavirus cases and deaths dayly in Poland",
               align = "lrrrr") %>% 
  kable_styling() %>% 
  kable_classic(full_width = FALSE)
```

## Covid-19 Vaccination Status

```{r}
Vaccine <- covid_tidy %>% 
  filter(date %in% as.Date(c("2020-12-28", "2020-12-31", "2021-01-31", "2021-02-28", 
                             "2021-03-31", "2021-04-29", "2021-05-31", "2021-06-30", 
                             "2021-07-31", "2021-08-30"))) %>% 
  select(date, total_vaccinations, total_vaccinations_per_hundred, 
         people_fully_vaccinated, people_fully_vaccinated_per_hundred) %>% 
datatable(escape = FALSE,
          filter = 'top',
          caption = htmltools::tags$caption("Vaccination Status monthly in Poland"),
          options = list(pageLength = 5,
                         class = "hover"),
          colnames = c("Date", "Total Vaccinations", "Total Vaccinations(100)", 
                       "People Vaccinated", "People Vaccinated(100)")
          )
Vaccine
```

# References

JJ Allaire, Rich Iannone, Alison Presmanes Hill and Yihui Xie (2021). distill: 'R Markdown' Format for Scientific and Technical Writing. R package version 1.2. https://CRAN.R-project.org/package=distill

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686

Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of Statistical Software, 40(3), 1-25. URL https://www.jstatsoft.org/v40/i03/.

Hao Zhu (2021). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.3.4. https://CRAN.R-project.org/package=kableExtra

Yihui Xie, Joe Cheng and Xianying Tan (2021). DT: A Wrapper of the JavaScript Library 'DataTables'. R package version 0.18. https://CRAN.R-project.org/package=DT

Ramnath Vaidyanathan, Kent Russell and Gareth Watts (2016). sparkline: 'jQuery' Sparkline 'htmlwidget'. R package version 2.0. https://CRAN.R-project.org/package=sparkline

WHO COVID-19 Dashboard. Geneva: World Health Organization, 2020. Available online: https://covid19.who.int/ 

Sharon Machlis (2018). How to add sparklines to R tables. InfoWorld. https://www.infoworld.com/article/3318222/how-to-add-sparklines-to-r-tables.html